Trajectory model of passive kelp dispersal

### 2020-04-23 thesis_chapter_3.Rmd
alterations to chapter 3 and updated output and template

---
title: 'Trajectory model of passive kelp dispersal'
author: "Ross Coppin"
output:
  html_document:
    df_print: paged
  word_document:
    reference_docx: markdown/thesis_template.docx
  pdf_document:
    fig_caption: yes
bibliography: references.bib
---

@@

There is a range of objects, both natural and anthropogenic, floating in the ocean but macroalgae are regarded as one of the most important passive dispersal mechanisms of marine taxa. Floating kelp as a passive dispersal mechanism for a range of marine taxa is well recognised and is sometimes referred to the tumble weed of the ocean [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008]. Although in recent years much of the focus of flotsam studies have shifted towards microplastics, considerable research has been performed investigating macroalgae as a passive dispersal mechanism in the ocean. Some macroalgae species are negatively buoyant and sink to the seafloor when detached from the substrate; usually through intense wave power. Other species of macroalgae have air-filled pneumatocysts or stipes, which allow the plants to reach the surface where light is more abundant. In turn, the positively buoyant pneumatocysts cause plants to float to the surface when dislodged from the substratum. The structure and number of pneumatocysts varies between species. For example species from the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other kelp species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, *Ecklonia maxima* have a single, large pneumatocysts [@dayton1985; @smith2002; @thiel2005b; @graiff2016; @batista2018]. In some cases the stipe itself is air-filled, such as with *E. maxima* and *E. radiata*. Although there have been reports of floating of Chlorophyta and some Rhodophyta species, Phaeophyceae species are the most commonly reported forms of floating algae. This is most likely because the green and red species reported floating are not actually positively buoyant, but instead are kept at the surface by gas trapped inbetween or in the thalli [@dromgoole1982; @back2000]. The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018] and the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018] have been the focus of much of the research regarding floating macroalgae and the ecological role of rafting. Comparatively, little research has been conducted on the dispersal characteristics of other Phaeophyceae species.

The raft-time or float-time, ie. the amount of time macroalgae floats at the surface before becoming negatively buoyant and sinks vary across species. Raft-time also referred to as raft longevity, is determined by epibiont load, macroalgal growth rate and the loss of kelp biomass over time through fragmentation. Epibiont load consists of epiphytic and bryozoan species which tend to grow more rapidly in higher light and temperature environments, such as the ocean surface [@thiel2005b; @thiel2005a; @graiff2016]. The increase in epibiont load reduces the buoyancy of floating macroalgal rafts over time eventually causing them to sink to the seafloor, while also potentially increasing drag [@hobday2000; @tala2013; @craw2018]. Fragmentation of kelp biomass over time may also reduce raft times further, while growth may counteract the effects of epibiont load and fragmentation provided the environmental conditions are favorable [@graiff2016; @macaya2016]. For instance a study by @rothausler2011 investigated the abiotic and biotic factors influencing raft time and dispersal potential of the giant kelp *M. pyrifera* by a combination of tethering experiments and field surveys. The results showed the physiological performance of kelp declined with increasing epiphyte biomass and that higher temperatures increased growth rate of epibionts and fragmentation of kelp. The authors concluded from the results of both the experiments and field surveys that *M. pyrifera* dispersal is dependent on low temperature and moderate irradiance conditions, with high temperatures and higher irradiance reducing overall raft-time and hence dispersal potential [@rothausler2011].

Many past authors have also taken various approaches to estimate not only the trajectory of macroalgal rafts but also the oceanographic influencers of macroalgae trajectory. Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005a, @thiel2005b; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011]. Although ocean currents are regarded as the primary influencer, the relative importance of wind versus surface current is still not known, although the role of wind has been recognised as important in several studies. For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M. pyrifera* in nearshore Monterrey Bay. The results showed that kelp rafts with little surface area exposed to the wind were largely driven by a combination of wind and wind waves, although the relative importance of wind and wind waves was not clear. In addition, the tagged kelp trajectories were more consistent with the formation of eddies during winter. Previous studies have identified wind as an important mechanism of dispersal in wind dominated ocean systems. For example the subAntartic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating kelp. Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013]. For instance, a study by @Fraser2011 on the rafting capabilities of *Durvillaea antarctica* used a combination of population genetics and relative age estimate of ‘goosebarnacles’ attached to the raft. The presence of goosebarnacles suggests a long raft time as these species have a slow growth rate, while the genetic analyses showed these species are able to raft up to $\sim$ 390km from their local origin. The authors suggested that wind and water-movement were the primary influences of trajectory, however, this was only inferred from the genetic results and local climatology data [@fraser2011].

@@

$$F_{c} = F_{kelphead} +  F_{holdfast} + F_{stipekeel}$$
where $F_{c}$ is overall seawater drag force, $F_{kelphead}$ is the estimated drag coefficient for the top portion of the plant (fronds, primary blade, pnuematocyst), $V_{holdfast}$ is the estimated drag coefficient for the holdfast, $V_{stipekeel}$ is the estimated drag coefficient for the volume of the stipe submerged. The assumption made is that the kelp-head and holdfast will not change in volume while floating and that the entire holdfast will be submerged. Provided the exposed volume of the kelp exposed to the air is know the submerged portion $V_{stipekeel}$ and overall mass can be calculated based on Archimedes law [@Andersson2017]:
$$V_{stipekeel} = V_{stipesail}\frac{p_{ice}}{p_w-p_{ice}}$$,

$$m = (V_{stipesail}+V_{stipekeel})p_{ice} =(1 + \frac{p_w}{p_w -p{ice}})V_{stipesail}p_{ice}$$
To calculate the volume of the stipe a idealised cyliderical form is used with height being subsituted for height such that:
$$V = \pi r^2 l$$

>>>>

---
title: 'Trajectory model of passive kelp dispersal'
author: "Ross Coppin"
output:
  word_document:
    reference_docx: markdown/thesis_template.docx
  html_document:
    df_print: paged
  pdf_document:
    fig_caption: yes
bibliography: references.bib
---

@@

There is a range of objects, both natural and anthropogenic, floating in the ocean but macroalgae are regarded as one of the most important passive dispersal mechanisms of marine taxa. Floating kelp as a passive dispersal mechanism for a range of marine taxa is well recognised and is sometimes referred to the tumble weed of the ocean [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008]. Although in recent years much of the focus of flotsam studies have shifted towards microplastics, considerable research has been performed investigating macroalgae as a passive dispersal mechanism in the ocean. Some macroalgae species are negatively buoyant and sink to the seafloor when detached from the substrate; usually through intense wave power. Other species of macroalgae have air-filled pneumatocysts or stipes, which allow the plants to reach the surface where light is more abundant. In turn, the positively buoyant pneumatocysts cause plants to float to the surface when dislodged from the substratum. The structure and number of pneumatocysts varies between species. For example species from the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other kelp species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, *Ecklonia maxima* have a single, large pneumatocysts [@dayton1985; @smith2002; @thiel2005; @graiff2016; @batista2018]. In some cases the stipe itself is air-filled, such as with *E. maxima* and *E. radiata*. Although there have been reports of floating of Chlorophyta and some Rhodophyta species, Phaeophyceae species are the most commonly reported forms of floating algae. This is most likely because the green and red species reported floating are not actually positively buoyant, but instead are kept at the surface by gas trapped inbetween or in the thalli [@dromgoole1982; @back2000]. The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018] and the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018] have been the focus of much of the research regarding floating macroalgae and the ecological role of rafting. Comparatively, little research has been conducted on the dispersal characteristics of other Phaeophyceae species.

The raft-time or float-time, ie. the amount of time macroalgae floats at the surface before becoming negatively buoyant and sinks vary across species. Raft-time also referred to as raft longevity, is determined by epibiont load, macroalgal growth rate and the loss of kelp biomass over time through fragmentation. Epibiont load consists of epiphytic and bryozoan species which tend to grow more rapidly in higher light and temperature environments, such as the ocean surface [@thiel2005; @thiel2005; @graiff2016]. The increase in epibiont load reduces the buoyancy of floating macroalgal rafts over time eventually causing them to sink to the seafloor, while also potentially increasing drag [@hobday2000; @tala2013; @craw2018]. Fragmentation of kelp biomass over time may also reduce raft times further, while growth may counteract the effects of epibiont load and fragmentation provided the environmental conditions are favorable [@graiff2016; @macaya2016]. For instance a study by @rothausler2011 investigated the abiotic and biotic factors influencing raft time and dispersal potential of the giant kelp *M. pyrifera* by a combination of tethering experiments and field surveys. The results showed the physiological performance of kelp declined with increasing epiphyte biomass and that higher temperatures increased growth rate of epibionts and fragmentation of kelp. The authors concluded from the results of both the experiments and field surveys that *M. pyrifera* dispersal is dependent on low temperature and moderate irradiance conditions, with high temperatures and higher irradiance reducing overall raft-time and hence dispersal potential [@rothausler2011].

Many past authors have also taken various approaches to estimate not only the trajectory of macroalgal rafts but also the oceanographic influencers of macroalgae trajectory. Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005, @thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011]. Although ocean currents are regarded as the primary influencer, the relative importance of wind versus surface current is still not known, although the role of wind has been recognised as important in several studies. For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M. pyrifera* in nearshore Monterrey Bay. The results showed that kelp rafts with little surface area exposed to the wind were largely driven by a combination of wind and wind waves, although the relative importance of wind and wind waves was not clear. In addition, the tagged kelp trajectories were more consistent with the formation of eddies during winter. Previous studies have identified wind as an important mechanism of dispersal in wind dominated ocean systems. For example the subAntartic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating kelp. Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013]. For instance, a study by @Fraser2011 on the rafting capabilities of *Durvillaea antarctica* used a combination of population genetics and relative age estimate of ‘goosebarnacles’ attached to the raft. The presence of goosebarnacles suggests a long raft time as these species have a slow growth rate, while the genetic analyses showed these species are able to raft up to $\sim$ 390km from their local origin. The authors suggested that wind and water-movement were the primary influences of trajectory, however, this was only inferred from the genetic results and local climatology data [@fraser2011].

@@

$$F_{c} = F_{kelphead} +  F_{holdfast} + F_{stipekeel}$$
where $F_{c}$ is overall seawater drag force, $F_{kelphead}$ is the estimated drag coefficient for the top portion of the plant (fronds, primary blade, pnuematocyst), $V_{holdfast}$ is the estimated drag coefficient for the holdfast, $V_{stipekeel}$ is the estimated drag coefficient for the volume of the stipe submerged. The assumption made is that the kelp-head and holdfast will not change in volume while floating and that the entire holdfast will be submerged. Provided the exposed volume of the kelp exposed to the air is know the submerged portion $V_{stipekeel}$ and overall mass can be calculated based on Archimedes law [@Andersson2017]:
$$V_{stipekeel} = V_{stipesail}\frac{p_{ice}}{p_w-p_{ice}}$$

and so,
$$m = (V_{stipesail}+V_{stipekeel})p_{ice} =(1 + \frac{p_w}{p_w -p{ice}})V_{stipesail}p_{ice}$$
To calculate the volume of the stipe a idealised cyliderical form is used with height being subsituted for height such that:
$$V = \pi r^2 l$$
